---
title: "US Election - County Analysis"
output: html_notebook
---


Load in required packages
```{r}
required_Packages = list("dplyr", "ggplot2", "maps", "grid", "gridExtra", "glmnet", "scales" ) 
for( package in required_Packages )
{
  if( !is.element(package, installed.packages()[,1]))
  {
    install.packages(package)
  }
  require(package, character.only = T)
}

rm(required_Packages)
```




Load the county data in and combined the tables.
```{r}
county_Data <- list()

## get voting data
{ Results_By_County_DF <- read.csv( "US_County_Level_Presidential_Results_12-16.csv" )
  county_Data[['voting_Data']] <- data.frame(  'fips' = Results_By_County_DF$FIPS, 
                                               'votes_dem_2016' = Results_By_County_DF$votes_dem_2016,
                                               'votes_gop_2016' = Results_By_County_DF$votes_gop_2016,
                                               'total_votes_2016' = Results_By_County_DF$total_votes_2016,
                                               'per_dem_2016' = Results_By_County_DF$per_dem_2016,
                                               'per_gop_2016' = Results_By_County_DF$per_gop_2016,
                                               'per_point_diff_2016' = Results_By_County_DF$per_point_diff_2016,
                                               'per_point_diff_2012' = Results_By_County_DF$per_point_diff_2012 )
  rm(Results_By_County_DF)
}

# get county variables field description
county_Data[['fieldDescription']] <- read.csv("county_facts_dictionary.csv")

# get county factors, rename, and delete unneeded varaibles
{
  county_Info <- read.csv( "county_facts.csv" )
  { # data conversion
      { # convert need variables to percentages
          cols = as.vector( county_Data$fieldDescription$column_name[county_Data$fieldDescription$Convert.To.Percent=='y' ] )
          county_Info[ , cols ] <-  county_Info[ , cols ] /100
      }
      
      { # convert needed variables to per capita 
          cols = as.vector( county_Data$fieldDescription$column_name[county_Data$fieldDescription$Convert.to.per.pop=='y' ] )
          county_Info[ , cols ] <-  county_Info[ , cols ] / county_Info[ , 'PST045214' ]
      }
    
      { # change column names
          if( sum(colnames(county_Info)[4:ncol(county_Info)] != county_Data$fieldDescription$column_name) >0 )
          {
            print("Fix name order")
          } else
          {
            colnames(county_Info)[4:ncol(county_Info)] = as.character(county_Data$fieldDescription$New.Name)
          }
      }
    
      {  # remove unwanted columns
          county_Info[, colnames(county_Info) == 'Delete']  <- NULL
      }
    
      county_Data[['county_Info']] <- county_Info
  }
}

rm(county_Info)
```


The US map function
```{r}
colors = list( rep.Red = '#E91D0E', dem.Blue = '#232066' )  

usMap <- function( df, fillName, customColor = c(colors$rep.Red, 'white', colors$dem.Blue), addCountyLine=F, legend_name=F, setMinMax=F )
{
  # join the fill values with the county long/lat data
  fips <- county.fips
  fips <- cbind( fips,  do.call( 'rbind', strsplit( as.vector(fips$polyname ), split = ',') ) )
  fips$polyname <- NULL
  colnames(fips) <- c("fips", "region", "subregion" )
  fips = full_join( fips, map_data("county"), by = c("region", "subregion")  )
  fips = left_join( df, fips, by=c('fips'))
  
  # sets the fill value range ( 0 is the center value )
  if( setMinMax == F )
  {
    maxValue <- max(0.001, max(fips[, fillName] ))
    minValue <- min(-0.001, min(fips[, fillName] ) )
  } else
  {
    minValue = min( -0.001, setMinMax[1] )
    maxValue = max( 0.001, setMinMax[2] )
  }
  
  # plot the map
  g <- ggplot(fips, aes(x=long, y=lat, group = group)) + 
  geom_polygon( aes_string( group = 'group', fill=fillName )) +
  scale_fill_gradientn( colours=customColor,
                        values=rescale(c(minValue,0,maxValue)), 
                        guide = "colorbar", 
                        limits=c(minValue,maxValue),
                        labels = percent ) +
  theme(axis.title=element_blank(),
        axis.text=element_blank(),
        axis.ticks=element_blank(),
        panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        panel.background = element_blank()) 
  
  # remove the legend
  if( legend_name == F)
  {
    g = g + theme(legend.position="none")
  }
  
  # add county lines to map
  if(addCountyLine)
    g = g +geom_path( data = map_data("county") , color = "grey", size = 0.3) 
  
  # add state lines to map
  g = g +geom_path( data = map_data("state") , colour = "black", size=1)
  print(g)
}

```






Look at the county vote data to determine cause of change
```{r}

# get voting data
{ 
  voting_Data <- county_Data$voting_Data[ !is.na(county_Data$voting_Data$per_point_diff_2012) , ]
  class <- rep(-1, nrow(voting_Data))
  class[voting_Data$per_point_diff_2016>0] <- 1
  voting_Data$change <- voting_Data$per_point_diff_2016 - voting_Data$per_point_diff_2012
}
```

Plot some maps
```{r}
# map of majority. red if rep won, blue if dem wond
usMap( cbind(voting_Data, 'class' = class) , 'class', addCountyLine=T )

# map of 2012 difference in percentages
usMap(voting_Data, 'per_point_diff_2012', addCountyLine=T, 
      setMinMax=c( min( c(voting_Data$per_point_diff_2012,voting_Data$per_point_diff_2016 )),  
                   max( c(voting_Data$per_point_diff_2012,voting_Data$per_point_diff_2016 )) ) )

# map of change from 2012 to 2016
usMap(voting_Data, 'change' )

# map of 2016 difference in percentages
usMap(voting_Data, 'per_point_diff_2016', addCountyLine=T, 
      setMinMax=c( min( c(voting_Data$per_point_diff_2012,voting_Data$per_point_diff_2016 )),  
                   max( c(voting_Data$per_point_diff_2012,voting_Data$per_point_diff_2016 )) ) )

# same as above but get the legend
usMap(voting_Data, 'per_point_diff_2016',legend_name="Percent Differences",
      setMinMax=c( min( c(voting_Data$per_point_diff_2012,voting_Data$per_point_diff_2016 )),  
                   max( c(voting_Data$per_point_diff_2012,voting_Data$per_point_diff_2016 )) ) )

```

county majoriy counts for 2012 and 2016
```{r}
sum(voting_Data$per_point_diff_2012>0) #dem2012  
sum(voting_Data$per_point_diff_2012<0) #rep2012 

sum(voting_Data$per_point_diff_2016>0) #dem2016 
sum(voting_Data$per_point_diff_2016<0) #rep2016  
```

look at changes between 2012 and 2016
```{r}
{
  sum( voting_Data$change<0 ) # count of decrease dem support
  sum( voting_Data$change>0 ) # count of increase in dem support
  
  # dem retained
  nrow( voting_Data[voting_Data$per_point_diff_2012>0,] )
  nrow( voting_Data[ (voting_Data$per_point_diff_2012>0 & voting_Data$per_point_diff_2016>0) ,] )
  # rep retained
  nrow( voting_Data[voting_Data$per_point_diff_2012<0,] )
  nrow( voting_Data[ (voting_Data$per_point_diff_2012<0 & voting_Data$per_point_diff_2016<0) ,] )
}

# histogram of the percent change for dem
{
  switchCounties <- voting_Data[(voting_Data$per_point_diff_2012>0 & voting_Data$per_point_diff_2016<0),]
  
  ggplot(switchCounties) + geom_histogram( aes( x=change ), color='black',fill = colors$dem.Blue, bins=15 ) + 
    xlab( "(2016 Percentage Difference) - (2012 Percentage Difference)") + ylab("Number of Counties") +
    scale_x_continuous(labels=percent) +
    ggtitle("Change in Percent Differences (Democrat minus Republican)" ) + 
    theme( axis.text.y = element_text(size = 12), 
           axis.text.x = element_text(size = 12),
           axis.title.y = element_text(size = 13), 
           axis.title.x = element_text(size = 13),
           plot.title = element_text(size = 18))
  
  rm(switchCounties)
}
```


Create dem_to_rep and dem_to_dem groups
```{r}
{
  dem_to_rep.index <- (voting_Data$per_point_diff_2012 > 0) &  (voting_Data$change < -.05) &  (voting_Data$per_point_diff_2016 < 0)
  dem_to_dem.index <- (voting_Data$per_point_diff_2012 > 0) &  (voting_Data$change > -.05) &  (voting_Data$per_point_diff_2016 > 0)
  sum(dem_to_dem.index)
  sum(dem_to_rep.index)
  dem_to_rep <- voting_Data[ dem_to_rep.index, ]
  dem_to_dem <- voting_Data[ dem_to_dem.index , ]
  
  # Map switched and remained counties  
  p <- rep( 0, nrow(voting_Data) )
  p[ dem_to_rep.index ] = -1
  p[ dem_to_dem.index ] = 1
  usMap( data.frame( voting_Data, p), 'p', addCountyLine =T )
  
  # Map switched and remained counties  with percent cahange
  p <- rep( 0, nrow(voting_Data) )
  p[ dem_to_rep.index ] = voting_Data$change[ dem_to_rep.index ]
  p[ dem_to_dem.index ] = voting_Data$change[ dem_to_dem.index ]
  usMap( data.frame( voting_Data, p), 'p', addCountyLine =T ,legend_name="add legend")
  usMap( data.frame( voting_Data, p), 'p', addCountyLine =T )
}

# clean up environment
{
  rm(voting_Data)
  rm(dem_to_rep.index)
  rm(dem_to_dem.index)
  rm(p)
  rm(class)
}

```





Create list objects for the two groups
```{r}
predictors <- colnames(county_Data$county_Info )[4:length(county_Data$county_Info)]

dem_to_dem <- list( data = left_join(dem_to_dem, county_Data$county_Info, by='fips' ) )
dem_to_dem[['diff']] <- dem_to_dem$data$per_point_diff_2016
dem_to_dem[['X']] <- dem_to_dem$data[, predictors] 
dem_to_dem[['n']] <- nrow(dem_to_dem$data)
dem_to_dem[['means']] <- colMeans( dem_to_dem$X )
dem_to_dem[['medians']] <-  apply(dem_to_dem$X, 2, median)
dem_to_dem[['var']] <-  diag( var( dem_to_dem$X ) )

dem_to_rep <- list( data = left_join(dem_to_rep, county_Data$county_Info, by='fips' ) )
dem_to_rep[['diff']] <- dem_to_rep$data$per_point_diff_2016
dem_to_rep[['X']] <- dem_to_rep$data[, predictors ]
dem_to_rep[['n']] <- nrow(dem_to_rep$data)
dem_to_rep[['means']] <- colMeans( dem_to_rep$X )
dem_to_rep[['medians']] <-apply(dem_to_rep$X, 2, median)
dem_to_rep[['var']] <-  diag( var( dem_to_rep$X ) )

```


fitted densities for the variables of the two groups
```{r}
for( pred in predictors)
{
  g <- ggplot() + geom_density( mapping = aes(x=dem_to_dem$data[,pred]), color = "blue" )  +
    geom_density( mapping = aes(x=dem_to_rep$data[,pred]  ), colour = "red") + xlab(pred) 
  print(g)
} 
```

Compare the means of the two groups ( switched - remained )
```{r}
diff <- list( mean =  dem_to_rep$means - dem_to_dem$means )
diff[['se']] <- numeric( length = length(colnames(dem_to_dem$X)))
names(diff$se) <- colnames(dem_to_dem$X)
diff[['p-value']] <- numeric( length = length(colnames(dem_to_dem$X)))
names(diff$`p-value`) <- colnames(dem_to_dem$X)

# calcualte SE and p-value for mean differences 
local( 
  for( pred in predictors )
  {
    if( grepl('%', pred))
    {
      p.dem <- dem_to_dem$means[pred]; q.dem = 1 - p.dem
      p.rep <- dem_to_rep$means[pred]; q.rep = 1 - p.rep
      se <- ( p.dem*q.dem / dem_to_rep$n + p.rep*q.rep / dem_to_dem$n )**0.5 
    } else
    {
      se <- ( dem_to_dem$var[pred] / dem_to_dem$n + dem_to_rep$var[pred] / dem_to_rep$n )**0.5
    }
    diff$se[pred] <<- se
    diff$`p-value`[pred] <<- 2* (1-pnorm(abs(diff$mean[pred]) / se) )
  }
)

# use bonferroni 
meanDiff <- sort( names( sort( diff$`p-value`[ diff$`p-value` < .05/length(colnames(dem_to_dem$X)) ])) )
print(sort( meanDiff))
```

Compare the median of the two groups ( switched - remained )
```{r}
# compare medians
median <- list( diff = apply(dem_to_rep$X, 2, median) - apply(dem_to_dem$X, 2, median)   )
median[['se']] <- numeric( length = length(colnames(dem_to_dem$X)))
median[['p-value']] <- numeric( length = length(colnames(dem_to_dem$X)))
names(median$`p-value`) <- colnames(dem_to_dem$X)

# calculate se and p-value of diff using bootstrap
local( 
  for( pred in predictors )
  {
    medians <- c()
    for( sim in c(1:200))
    {
      diffInMed <- median( sample(x=dem_to_rep$data[, pred], size =150, replace = T )) -
        median( sample(x=dem_to_dem$data[, pred], size =150, replace = T ))  
      medians <- c( medians, diffInMed)
    }
    median$se[pred] <<- var(medians)**0.5
    median$`p-value`[pred] <<- pnorm( q = abs(mean(medians) ), sd = var(medians)**0.5, lower.tail=FALSE  )
  }
)

# use bonferroni p-value
medDifff <- sort( names( sort( median$`p-value`[ median$`p-value` < .05/length(colnames(dem_to_dem$X)) ])) )
print(sort( medDifff))

```

plot mean and medium distributions ( switched - remained )
```{r}
for( pred in predictors)
{
  avg <- diff$mean[pred]
  se <- diff$se[pred]
  x = seq( from=avg-3*se, to=avg+3*se, length.out =100 )
  g1 <- ggplot() + geom_line( mapping = aes(x=x, y=dnorm(x, mean = avg, sd=se   )), color = "blue" )  + xlab(pred) + ggtitle("Mean") 

  
  avg <- median$diff[pred]
  se <- median$se[pred]
  x = seq( from=avg-3*se, to=avg+3*se, length.out =100 )
  g2 <- ggplot() + geom_line( mapping = aes(x=x, y=dnorm(x, mean = avg, sd=se   )), color = "blue" )  + xlab(pred) + ggtitle("Median") 
  
  if(se==0){next }
  print(grid.arrange(g1, g2, nrow = 1))
}
```


From above analysis lets create some group characteristics
```{r}
predictorsGrouping <- list()

predictorsGrouping[["Population"]]  <- c("Population", 
                                         "Population Change (%)", 
                                         "House: Multi-unit housing (%)", 
                                         "Population density",
                                         "Pop: Living in same place for a year (%)"  )

predictorsGrouping[["Race"]] <- c('Race: White (%)', 
                                  'Race: Black (%)', 
                                  'Race: Native (%)', 
                                  'Race: Asian (%)',
                                  'Race: Hawaiian (%)',
                                  'Raced: Mixed (%)',
                                  'Raced: Latino (%)',
                                  "Pop: Foreign Born (%)",
                                  "Non English Language spoken (%)" )

predictorsGrouping[["Age"]] <- c('Age: < 5 (%)', 
                                 'Age: < 18 (%)', 
                                 'Age:  > 65 (%)'  )

predictorsGrouping[["Education"]] <- c('High School grad (%)', 
                                       'Bachelors (%)'  )

predictorsGrouping[["Housing"]] <- c('House: Homeownership Rate (%)', 
                                     'House: Median Value', 
                                     'House: Person per household'  )

predictorsGrouping[["income"]] <- c('Income: per capital', 
                                    'Income: household income' 
                                     )

predictorsGrouping[["Economic"]] <- c('Porverty rate (%)', 
                                      "Economy: nonfarm employment", 
                                      "Economy: num of firms", 
                                      "Economy: Retail sales", 
                                      "Economy: wholesaler sales", 
                                      "Economy: Building permits",
                                      "Economy: Acc. And food servces sales",
                                      "Economy: Manufacturers shipments")
```


Now lets calculate the means and medians for those groups
```{r}

# print the mean and mediam for the two groups
for( group in names(predictorsGrouping ))
{
  print( paste(rep("=", 100), sep ='', collapse = ''))
  print(group)
  df= data.frame(  round(dem_to_rep$means[ predictorsGrouping[[group]]  ],5) ) 
  df = cbind( df,  data.frame(  round(dem_to_rep$medians[ predictorsGrouping[[group]]  ],5) ) )
  df = cbind( df,  data.frame(  round(dem_to_dem$means[ predictorsGrouping[[group]]  ],5) ) )
  df = cbind( df,  data.frame(  round(dem_to_dem$medians[ predictorsGrouping[[group]]  ],5) ) )
  
  
  colnames(df) <- c( paste( group, " - mean: dem_to_rep"),
                    paste( group, " - median: dem_to_rep"),
                    paste( group, " - mean: dem_to_dem"),
                    paste( group, " - median: dem_to_dem"))
  print(df) 
}

# need to do age sepeately
print( paste(rep("-", 100), sep ='', collapse = ''))
print( "mean: dem_to_rep")
round(dem_to_rep$means[ predictorsGrouping$Age],5)
mean( 1-  as.vector(rowSums(dem_to_rep$X[, predictorsGrouping[["Age"]]])  ) )

print( paste(rep("-", 100), sep ='', collapse = ''))
print( "mediam: dem_to_rep")
round(dem_to_rep$medians[ predictorsGrouping$Age],5)
median( 1-  as.vector(rowSums(dem_to_rep$X[, predictorsGrouping[["Age"]]])  ) )

print( paste(rep("-", 100), sep ='', collapse = ''))
print( "mean: dem_to_dem")
round(dem_to_dem$means[ predictorsGrouping$Age],5)
mean( 1-  as.vector(rowSums(dem_to_dem$X[, predictorsGrouping[["Age"]]])  ) )

print( paste(rep("-", 100), sep ='', collapse = ''))
print( "mediam: dem_to_dem")
round(dem_to_dem$medians[ predictorsGrouping$Age],5)
median( 1-  as.vector(rowSums(dem_to_dem$X[, predictorsGrouping[["Age"]]])  ) )

```


fit a logistic model.
Will use lasso to find important variabels
```{r}
{ # get the data
  x <- rbind(  dem_to_dem$X, dem_to_rep$X  )
  x$`Land Area` <- NULL
  y <-  data.frame( 'switched'=as.factor(c( rep(0, nrow(dem_to_dem$X)), rep(1, nrow(dem_to_rep$X))  ) ))
  df= cbind(x,y)
}

{ # get training and test dta
  set.seed(1)
  testsize = 200
  training_Indexes <- sample( c(1:nrow(x)), nrow(x)-testsize )
  testingIndex = setdiff( c(1:nrow(x)), training_Indexes)
  df_train <- df[training_Indexes, ]
}

{ # fit using lasso
  cv.out <- cv.glmnet( as.matrix(x),as.matrix(y),alpha=1,family='binomial' )
  plot(cv.out)
  lambda_min <- cv.out$lambda.min
  lambda_1se <- cv.out$lambda.1se
  coef(cv.out,s=lambda_1se)
}

# accuracy on test data
sum( round( predict(cv.out, as.matrix(x[ testingIndex , ]) ,type="response") ) == y[testingIndex,]) / testsize


# Also fitted models using stepwise
{ # use stepwise 
  # model 1
  model.min <- glm( switched ~1, data = df_train, family = 'binomial' )
  model.max <- formula(lm(switched~. ,df_train))
  fittedModel1 = step(model.min, direction='both', scope=model.max)
  sum( round( predict(fittedModel1, x[ testingIndex , ],type="response") ) == y[testingIndex,]) / testsize
  summary(fittedModel1)
  
    # model 2
  model.min <- glm( switched ~1+ `Bachelors (%)`, data = df_train, family = 'binomial' )
  model.max <- formula(lm(switched~. ,df_train))
  fittedModel2 = step(model.min, direction='both', scope=model.max)
  sum( round( predict(fittedModel2, x[ testingIndex , ],type="response") ) == y[testingIndex,]) / testsize
  summary(fittedModel2)
  
  # model 3
  model.min <- glm( switched ~1+`Race: White (%)`, data = df_train, family = 'binomial' )
  model.max <- formula(lm(switched~. ,df_train))
  fittedModel3 = step(model.min, direction='both', scope=model.max)
  sum( round( predict(fittedModel3, x[ testingIndex , ],type="response") ) == y[testingIndex,]) / testsize
  summary(fittedModel3)
  
  # model 4
  model.min <- glm( switched ~1+`Race: White (%)` + `Bachelors (%)`, data = df_train, family = 'binomial' )
  model.max <- formula(lm(switched~. ,df_train))
  fittedModel4 = step(model.min, direction='both', scope=model.max)
  sum( round( predict(fittedModel4, x[ testingIndex , ],type="response") ) == y[testingIndex,]) / testsize
  summary(fittedModel4)
}


```


Lets look more closely at white/bachelor relationship
```{r}
{ # Scatter plot of bachelor vs. white with county gouping
ggplot() + geom_point(data=df, aes( x = `Race: White (%)`, y =`Bachelors (%)`, colour=switched  )) +
  scale_colour_manual(  values = c(colors$dem.Blue, colors$rep.Red), labels=c("Remained", "Switched"), name="County Group") +
  scale_y_continuous(labels=percent) + scale_x_continuous(labels=percent) +
  ylab("Bachelors (%)") + xlab("White Population (%)") +
  ggtitle("Counties by Education and Race") +
  theme( axis.text.x = element_text(size = 12), 
         axis.text.y = element_text(size = 12),
         axis.title.y = element_text(size = 14),
         axis.title.x = element_text(size = 14),
         plot.title = element_text(size = 20)  )
}    


{ # fit logistic with white and bachelor and plot probabilities
  
  { # fit using lasso only using white and bachelor
    cv.out2 <- cv.glmnet( as.matrix(df_train[, c("Race: White (%)", "Bachelors (%)")] ),as.matrix(df_train$switched),alpha=1,family='binomial' )
    coef = coef(cv.out2,s=lambda_1se)
    sum( round( predict(cv.out2, as.matrix(x[testingIndex, c("Race: White (%)", "Bachelors (%)")]) ,type="response") ) == y[testingIndex,]) / testsize
  }
  
  
  p <- function( white )
  {
    logOdd <- coef['(Intercept)', 1] + coef['Race: White (%)', 1] * white + coef['Bachelors (%)', 1] * bachelor_per
    return( 1/(1+exp(-logOdd)) )
  }
  bachelor_per <- seq(0.1, 0.5, length.out=200) 
  ggplot() +geom_line( aes(x=bachelor_per, y = p(0.2), colour = 'white: 20%'    )) +
            geom_line( aes(x=bachelor_per, y = p(0.5), colour = 'white: 50%'  )) +
            geom_line( aes(x=bachelor_per, y = p(0.75), colour = 'white: 75%'  )) +
            scale_color_manual("", breaks=c('white: 20%','white: 50%','white: 75%' ), values=c("red", "blue","green")) +
            scale_y_continuous(labels=percent) + scale_x_continuous(labels=percent) +
            xlab("Bachelors (%)") + ylab("Estimated Probability") +
            ggtitle("Probability of Switch") +
            theme( axis.text.x = element_text(size = 12), 
                   axis.text.y = element_text(size = 12),
                   axis.title.y = element_text(size = 14),
                   axis.title.x = element_text(size = 14),
                   plot.title = element_text(size = 20)  )
}

# correlations 
colnames(x)[cor(x[,"Bachelors (%)"], x) < -.2]
colnames(x)[cor(x[,"Race: White (%)"], x) > .2]


```
